---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Marcus Minko"
date: "02-08-2022"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
dn <- dennys
lq <- laquinta
```

### Exercise 1

```{r}
dn_ak <- dn %>%
  filter(state == "AK")
nrow(dn_ak)
```


### Exercise 2

```{r}
lq_ak <- lq %>%
  filter(state == "AK")
nrow(lq_ak)
```
### Exercise 2

``` {r data-join}

dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak
```
#There are 6 observations and 11 variables (address.x, city.x, state, zip.x, longitude.x, latitude.x, address.y, city.y, zip.y, langitude.y, latitude.y)


### Exercise 3
```{r haversine}

haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```
...

### Exercise 4
```{r distance-mutate}
dn_lq_ak <- dn_lq_ak %>% 
    mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))
```
...

### Exercise 5
```  {r closest-distance}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

dn_lq_ak_mindist %>% 
     summarize(closest)

dn_lq_ak_mindist %>% 
     summarize(mean(closest))
```

### Exercise 6
``` {r viz}
dn_lq_ak_mindist %>% 
            ggplot(mapping = aes(x = closest, 
                                 y = address.x)) +
            geom_col() + 
            labs(
              title = "Distance between Denny's and La Quinta", subtitle = "Alaska", 
              x = "Distance to nearest La Quinta (km)", 
              y = "Denny's location")
```
...


### Exercise 6
``` {r NC}

#filter NC dennys
dn_nc <- dn %>%
  filter(state == "NC")

#filter NC laquinta
lq_nc <- lq %>%
  filter(state == "NC")

# Join data
dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")
dn_lq_nc

# Mutate distance var
dn_lq_nc <- dn_lq_nc %>% 
    mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))

# Min distance
dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

dn_lq_nc_mindist %>% 
     summarize(closest)

dn_lq_nc_mindist %>% 
     summarize(mean(closest))

# Visualization

dn_lq_nc_mindist %>% 
            ggplot(mapping = aes(x = closest, 
                                 y = address.x)) +
            geom_col() + 
            labs(
              title = "Distance between Denny's and La Quinta", subtitle = "NC", 
              x = "Distance to nearest La Quinta (km)", 
              y = "Denny's location")

dn_lq_nc_mindist %>% 
               ggplot(mapping = aes(x = closest)) +
               geom_histogram(binwidth = 5) + 
            labs(
              title = "Distance between Denny's and La Quinta", subtitle = "North Carolina", 
              x = "Distance from Denny's to nearest La Quinta (km)", 
              y = "Number of Denny's")
```
...

### Exercise 7
``` {r TX}

#filter TX dennys
dn_tx <- dn %>%
  filter(state == "TX")

#filter TX laquinta
lq_tx <- lq %>%
  filter(state == "TX")

# Join data
dn_lq_tx <- full_join(dn_tx, lq_tx, by = "state")
dn_lq_tx

# Mutate distance var
dn_lq_tx <- dn_lq_tx %>% 
    mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))

# Min distance
dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

dn_lq_tx_mindist %>% 
     summarize(mean(closest))

# Visualization

dn_lq_tx_mindist %>% 
               ggplot(mapping = aes(x = closest)) +
               geom_histogram(binwidth = 5) + 
            labs(
              title = "Distance between Denny's and La Quinta", subtitle = "Texas", 
              x = "Distance from Denny's to nearest La Quinta (km)", 
              y = "Number of Denny's")
```
### Exercise 6
``` {r GA}

#filter GA dennys
dn_ga <- dn %>%
  filter(state == "GA")

#filter GA laquinta
lq_ga <- lq %>%
  filter(state == "GA")

# Join data
dn_lq_ga <- full_join(dn_ga, lq_ga, by = "state")
dn_lq_ga

# Mutate distance var
dn_lq_ga <- dn_lq_ga %>% 
    mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))

# Min distance
dn_lq_ga_mindist <- dn_lq_ga %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

dn_lq_ga_mindist %>% 
     summarize(mean(closest))

# Visualization

dn_lq_ga_mindist %>% 
            ggplot(mapping = aes(x = closest, 
                                 y = address.x)) +
            geom_col() + 
            labs(
              title = "Distance between Denny's and La Quinta", subtitle = "Georgia", 
              x = "Distance to nearest La Quinta (km)", 
              y = "Denny's location")

dn_lq_ga_mindist %>% 
               ggplot(mapping = aes(x = closest)) +
               geom_histogram(binwidth = 5) + 
            labs(
              title = "Distance between Denny's and La Quinta", subtitle = "Georgia", 
              x = "Distance from Denny's to nearest La Quinta (km)", 
              y = "Number of Denny's")

dn_lq_ak_mindist %>% 
    summarize(mean(closest))
dn_lq_nc_mindist %>% 
    summarize(mean(closest))
dn_lq_tx_mindist %>% 
    summarize(mean(closest))
dn_lq_ga_mindist %>% 
    summarize(mean(closest))
```
Hedberg's observation is probably most true for TX. Even though AK has a lower mean distance, there are so few observations that it probably shouldn't be considered.
